The role of advection versus diffusion in intraventricular transport

### 2025-02-04 paper/main.tex
clean Appendix

\section*{Methods}
\subsection*{Image-based computational geometries of zebrafish brain ventricles}
To model flow and transport in zebrafish larval brain ventricles, we generated a
3D representation of the brain ventricular walls using confocal imaging data of a zebrafish embryo
injected intraventricularly with a 70 kDa dye at 2 dpf~\cite{Olstad2019CiliaryDevelopment}. 
From the imaging data, we constructed a computational mesh of the ventricles (\Cref{fig:figure1}d).
The interior volume was meshed with fTetWild~\cite{Hu2020FastWild}. 
The \emph{standard} mesh consisted of 132 134 tetrahedral cells and had a
maximal (minimal) edge length of 14.9 $\mu$m (4.68 $\mu$m). 
At 2 dpf, the ventricular system consists mainly of 3 cavities: the telencephalic (anterior),
the di-/mesencephalic (middle) and the rhombencephalic (posterior) ventricles, which are connected by ducts. 
Parts of the ventricular surfaces were marked to be lined with
motile cilia~\cite{Olstad2019CiliaryDevelopment} (\Cref{fig:figure1}d, magenta markers).
\begin{figure}%[hbt!]
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{\lyng{Experimenting with velocity colorbars, thoughts on the
    two options? Or have two separate as it is now?}}
    \label{fig:figure1}
\end{figure}

@@

\subsection*{Computational CSF dynamics}
The beating motion of the cilia generates CSF flow with steady rotational
structures~\cite{Olstad2019CiliaryDevelopment}
within the brain ventricles (\Cref{fig:figure1}a--c).
In addition, cardiac pulsations induces pulsatile CSF flow.
We model this flow of CSF in the brain ventricles by the time-dependent,
incompressible Stokes equations which read as follows.
Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$
and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,\\
      \nabla \cdot \uu &= 0
      \quad  \text{in } \Omega,
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$
and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. 
Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions.
Furthermore, $\rho=1000 \ \mathrm{kg/m^3}$ is the CSF density
and $\mu=6.97\times 10^{-4} \ \mathrm{Pa\cdot s}$ is the dynamic viscosity
of the CSF~\cite{Bloomfield1998EffectsFluid}.
We introduce the tangential traction $\bsigpar$ as the
tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$,
and $P_{\nn}(\rr)$ is the tangential projection of a vector $\rr$ onto $\Gamma$.
We consider three types of boundary conditions:
(i) \emph{tangential traction} (no flow normal to the boundary
$\uu \cdot \nn = 0$ with a force $\btau$ applied tangentially on the boundary $\bsigpar = \btau$),
(ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in
the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$),
and (iii) \emph{free slip} (no flow normal to the boundary
$\uu \cdot \nn = 0$ and no tangential forces $\bsigpar = {\bf 0}$).
Unless noted otherwise, the free-slip condition is applied on the
boundary. We assume that the system starts at rest: $\uu(\xx, t=0) = \mathbf{0}$.

To study the impact of the motile cilia and the cardiac pulsations on the CSF flow,
we consider the following computational flow scenarios,
each predicting a CSF flow velocity $\uu$ and CSF pressure $p$.
\begin{itemize}
    \item A \emph{baseline} flow model representing a wild-type zebrafish including 
    contributions both from the cilia forces and the cardiac pulsations.
    In this scenario, we apply the tangential traction $\btau$ on the cilia boundary
    $\Gc$ and prescribe the normal pressure $\Tilde{p}$ on the anterior/posterior boundary $\Gp$. % Model 0

    \item A \emph{cilia-only/no-cardiac} flow model including contributions from the
    cilia forces but no cardiac pulsations.
    Here, we apply the tangential traction $\btau$ on the cilia boundary $\Gc$.
    In the absence of cardiac flow, the anterior/posterior boundary $\Gp=\emptyset$. % Model I

    \item A \emph{cardiac-only/no-cilia} flow model including contributions from the
    cardiac pulsations but no cilia forces.
    Here, we set the tangential traction to zero on the cilia boundary
    $\Gc$, while prescribing the normal pressure $\Tilde{p}$ on the
    anterior/posterior boundary $\Gp$.  % Model II
\end{itemize}

\subsection*{Cilia-driven flow}
On the ciliated regions of the ventricle walls $\Gc$ (\Cref{fig:figure1}d, magenta markers),
we impose a constant-in-time tangential traction $\bsigpar=\btau$ to represent the net forces
of the cilia acting on the CSF.
We set $\btau(\xx, \rr) = \tau \lambda(\xx) P_{\nn}(\rr)$, where the function $\lambda(\xx)$
and the sign of the vector $\rr=\pm(1, 0, 1)$ varies depending on the
cilia region (see~\Cref{subsec:appendixA1}).
Both $\rr$ and the value $\tau=6.5\times 10^{-4}$ Pa were chosen by 
model calibration using experimental data~\cite{Olstad2019CiliaryDevelopment}.
The data indicate speeds proximate to the ventricular walls of $27.4 \pm 5.4$ \textmu m/s and
$2.6 \pm 0.6$ \textmu m/s in the dorsal and ventral regions of the middle ventricle, respectively.
In the anterior ventricle, the speeds were estimated at $4.7 \pm 1.3$ \textmu m/s.
We calibrated $\tau$ heuristically by comparing these experimental velocity values
with the simulated, maximum velocity magnitude.

\subsection*{Flow induced by cardiac pulsations}\label{subsec:pressure_bcs}
To drive a cardiac-induced pulsatile flow inside the brain ventricles, we set $\Tilde{p}(t)=0$ and
$\Tilde{p}(t)=-A\sin{\omega t}$ on anterior and posterior parts, respectively, of the
boundary $\Gp$~(\Cref{fig:figure1}d, blue markers).
Based on a measured cardiac frequency of $f=2.22$ Hz~\cite{Olstad2019CiliaryDevelopment},
the angular cardiac frequency $\omega=2\pi f=6.97$ rad/s.
The amplitude $A=1.5\times 10^{-3}$ Pa determines the magnitude of the pulsatile flow,
and was calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}.
Similar to $\tau$, the amplitude $A$ was chosen by comparing experimental and simulated velocities.
Particle tracking data of pulsatile motion was used to determine the contributions of the
cardiac pulsations to the velocity magnitude,
as compared to the effect of cilia motion on the flow (see~\Cref{sec:appC_particle_transport}).

@@

\subsection*{Generation of \emph{Tg(5xuas:Signal-Dendra2)\textsuperscript{nw21Tg}} line}
\lyng{Needs a more understandable title.}
The open reading frame (ORF) of the neuropeptide y (npy) signal peptide
(\emph{npy}: ENSDARG00000036222, Q1LW93) was fused with the N-terminal
of zebrafish codon optimized Dendra2 DNA sequence.
The DNA sequence was synthesized with EcoR1 enzyme sites at the 5’ and
3’ ends (GenScript Biotech), and inserted into the
pT2MUASMCS vector~\cite{Asakawa2008GeneticZebrafish}
through restriction enzyme cloning (GenScript Biotech).

@@

\subsection*{Intraventricular injection of microbeads and particle tracking}
To recover the flow fields shown in~\Cref{fig:figure1}b, c, we carried out
particle tracking in a larval zebrafish injected with fluorescent beads.
As described in Olstad~\emph{et al.}~\cite{Olstad2019CiliaryDevelopment},
anesthetized 2 dpf zebrafish were injected with 1 nl of a mixture containing 0.1\% w/v fluorescent beads
(SPHERO Fluorescent Yellow Particles 1\% w/v, F = 0.16 mm) diluted
in 7.5 mg/ml 70 kDa rhodamine B isothiocyanate-dextran (RITCdextran; Sigma-Aldrich, R9379)
dissolved in artificial CSF. Artificial CSF composition was as follows:
124 mM NaCl, 22 mM D-(+)-Glucose, 2.0 mM KCl, 1.6 mM MgSO\textsubscript{4} $\cdot$ 7 H\textsubscript{2}O,
1.3 mM KH\textsubscript{2}PO\textsubscript{4}, 24 mM NaHCO\textsubscript{3},
2.0 mM CaCl\textsubscript{2} $\cdot$ 2 H\textsubscript{2}O.
The needles used for the injections were pulled with a
Sutter Instrument Co. Model P-2000 from thin-walled glass
capillaries (1.00 mm; VWR) and cut open with a forceps.
A volume of 1 nl of the solution was injected with a pressure injector
(Eppendorf Femtojet 4i) in the rostral rhombencephalic ventricle.
The pressure and time were calibrated for each needle using a 0.01 mm calibration slide.

Following injection, the zebrafish was directly placed under the confocal
microscope and 1200 images were acquired at a frequency of 13.1 Hz.
A single optical section was obtained. Particle tracking was done using
TrackMate~\cite{Tinevez2017TrackMate:Tracking} in Fiji/ImageJ~\cite{Schindelin2012Fiji:Analysis}
and plotted using MATLAB, as described in Jeong~\emph{et al.}~\cite{Jeong2022MeasurementTelencephalon}.
The parameters for TrackMate were as follows.
The DoG detector was used with the settings: threshold: 2.0, with
median filtering; radius: 2.0, with subpixel localization.
Next, the Simple LAP tracker was used with settings: max frame gap: 2;
alternative linking cost factor: 1.05; linking max distance: 5.0; gap closing max distance: 2.0;
splitting max distance: 15.0; allow gap closing: true; allow track splitting:
false; allow track merging: false; merging max distance: 15.0; cutoff percentile: 0.9.
Only the particle tracks with at least 30 data points, while simultaneously
covering a distance of at least 8.8 $\mu$m, were plotted in~\Cref{fig:figure1}b, c.

@@

We seek to model the compartimentalized and pulsatile CSF flow observed \emph{in vivo}~(\Cref{fig:figure1}b--c).
To characterize CSF flow patterns predicted by
computational fluid dynamics within the brain ventricles, we begin by
considering the simulated CSF velocity and pressure when only considering
cilia motion, disregarding cardiac pressure pulsations. The
bidirectional traction induced by the motile cilia
inherently leads to a partial compartmentalization of the ventricular
system, with large-scale vortex structures in the anterior and middle
ventricles~(\Cref{fig:figure1}e). The highest velocities ($26.5$ \textmu m/s)
occur in the vicinity of the dorsal cilia in the middle ventricle.

To examine the effect of cardiac pressure pulsations,
we next consider a no-cilia model variation. In this scenario, the simulated CSF
flow is pulsatile as a result of the sinusoidal pressure gradient (\Cref{fig:figure1}g),
with a peak velocity magnitude of $4.8$ \textmu m/s. At each time instant, the flow
is unidirectional, with CSF initially flowing rostrocaudally (positive $x-$axis) before changing
direction to caudorostral in the middle of the cardiac cycle.

We now consider the baseline model, simulating flow driven by both cilia
and cardiac pulsations. We observe \fixme{persistent} flow features,
with vortex-structures in the anterior and middle ventricles
and directional flow in the posterior ventricle (\Cref{fig:figure1}e,
\fixme{Supplementary video 1}). These characteristic patterns suggest
that the flow in the anterior and middle ventricles are dominated by
the cilia contribution, while the cardiac pressure pulsations dominate
within the posterior ventricle. Over one cardiac cycle, the CSF flow
speed reaches up to 27.7 \textmu m/s, dorsally in the posterior part
of the middle ventricle. This peak is attained at the time instant
coinciding with the time of peak caudal cardiac pulsatile flow, as a
result of both the traction induced by the cilia and the pulsatile
pressure-driven flow being aligned with the rostrocaudal 
axis. Using the observed maximum velocity magnitude and choosing the height of
the middle ventricle ($110$ \textmu m) as a characteristic length
scale, yields a Reynolds number $\mathrm{Re} \approx 0.004$,
justifying the assumption of Stokes flow. In summary, the computational
CSF flow model reproduces the flow features observed \emph{in vivo}.

@@

To quantify the transport of solute, we plot the profile of the mean
concentrations $\cbar_i(t)$ as functions of time
(\cref{fig:figure2_compare_diffusion_coefficients}\fixme{b--d}) and
the times $\hat{t}$ when the mean concentrations $\cbar_i$ first exceed
a threshold value $\hat{c}$ (\cref{fig:figure2_compare_diffusion_coefficients}\fixme{f})
in 6 regions of interest (ROIs) (\cref{fig:figure2_compare_diffusion_coefficients}\fixme{e}).
The dynamics at the site of photoconversion (ROI 1)
is logarithmic and follows from imposing \eqref{eq:photoconversion_curve}.
For ROIs 2--6, we observe faster kinetics with higher diffusion coefficients
(\Cref{fig:figure2_compare_diffusion_coefficients}\fixme{b--d}),
which results in smaller time to threshold
(\Cref{fig:figure2_compare_diffusion_coefficients}\fixme{f}).
For all three diffusion coefficients, the dynamics of ROI 5
is significantly delayed compared to ROI 2--4,
and ROI 6 is the slowest. Notably, for $D_1$ the mean concentration
in ROI 6 is close to zero at the end of simulation.
To quantify the balance of advection versus diffusion as transport mehcanisms,
we calculate the Péclet numbers $\mathrm{Pe}_1=519$,
$\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for $D_1$, $D_2$ and $D_3$, respectively.
Since $\mathrm{Pe}>1$, this indicates that advection is the
dominant transport mechanism on a \fixme{global} scale.

@@

\subsection*{Computational model assumptions and boundary conditions}
To model CSF, we employed slip boundary conditions at the ventricular walls,
contrasting the classical choice of a noslip boundary condition.
The use of slip boundary conditions is motivated by experimental observations of the
flow field in zebrafish embryo brain ventricles~\cite{Olstad2019CiliaryDevelopment}.
It has been observed, both previously and in this work, that the fluid velocity increases towards the wall~(\Cref{fig:figure1}c).
Using slip boundary conditions allows for this behavior.
A couple of previous continuum model simulation studies of cilia-mediated flow~\cite{Siyahhan2014FlowVentricles, Thouvenin2020OriginCanal}
have modeled cilia as body forces. We chose to model the cilia as a force exerted tangentially on the boundary.
First, we deem this a good representation of the real nature of cilia,
since they are in reality microscale structures confined to the ventricular walls that
generate flow through collective motion. Second, it avoids the need of excessive mesh
refinement near the ciliated boundaries in the case that a
body force were to be applied in a $\sim$ 5 $\mu$m~\cite{Salman2022ComputationalEmbryo} deep cilia layer.

@@

\lyng{HH: Add a paragraph discussing the advection-diffusion equation discretization with DG elements?}

@@

\textbf{Only-cilia/no-cardiac model.} This model version has cilia forces
as the only mechanism driving flow. We consider the
Stokes equations~\eqref{eq:appendix_stokes_eqs} with the slip boundary conditions
\begin{subequations}
    \begin{alignat}{2}
        \uu\cdot\nn &= 0 \quad &&\mathrm{on} \ \Gamma=\Gc\cup\Gs, \label{eq:modelA_slip_BC_impermeability} \\
        %\hat{\sigma} - (\hat{\sigma}\cdot\nn)\nn &= \tau P_{\nn}(\rr) \quad &&\mathrm{on} \ \Gamma_s,
        \bsigpar &= \btau \quad &&\mathrm{on} \ \Gc,
        \label{eq:modelA_slip_BC_shear} \\
        \bsigpar &= \mathbf{0} \quad &&\mathrm{on} \ \Gs,\label{eq:modelA_full_slip_BC}
    \end{alignat}\label{eq:modelA_slip_BC}%
\end{subequations}%
on the boundary $\Gamma=\Gamma_c\cup\Gamma_s$. The boundary $\Gamma_s$ is a free-slip boundary, where we impose the free-slip condition~\eqref{eq:modelA_full_slip_BC}. The first condition~\eqref{eq:modelA_slip_BC_impermeability} is the impermeability condition, enforcing no flow normal to the boundary, whereas~\eqref{eq:modelA_slip_BC_shear} imposes a stress tangential to the boundary. These tangential stresses are applied to represent the forces that the collective cilia motion exerts on the cerebrospinal fluid.

@@

in the weak form~\eqref{eq:modelA_weak_form1} of the Stokes equations. We decompose the traction vector $\hat{\bsig}$ into a normal and tangential component:

@@

Inserting the above into~\eqref{eq:boundary_integral_BC1}, combined with application of the tangential traction condition~\eqref{eq:modelA_slip_BC_shear} and the free-slip condition~\eqref{eq:modelA_full_slip_BC}, yields

>>>>

\section*{Methods}
\subsection*{Image-based computational geometries of zebrafish brain ventricles}
To model flow and transport in zebrafish larval brain ventricles, we generated a
3D representation of the brain ventricular walls using confocal imaging data of a zebrafish embryo
injected intraventricularly with a 70 kDa dye at 2 dpf~\cite{Olstad2019CiliaryDevelopment}. 
From the imaging data, we constructed a computational mesh of the ventricles (\Cref{fig:fig1}d).
The interior volume was meshed with fTetWild~\cite{Hu2020FastWild}. 
The \emph{standard} mesh consisted of 132 134 tetrahedral cells and had a
maximal (minimal) edge length of 14.9 $\mu$m (4.68 $\mu$m). 
At 2 dpf, the ventricular system consists mainly of 3 cavities: the telencephalic (anterior),
the di-/mesencephalic (middle) and the rhombencephalic (posterior) ventricles, which are connected by ducts. 
Parts of the ventricular surfaces were marked to be lined with
motile cilia~\cite{Olstad2019CiliaryDevelopment} (\Cref{fig:fig1}d, magenta markers).
\begin{figure}%[hbt!]
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{\fixme{Reorder panels.} 
    \textbf{a.} Schematic illustration of the two cerebrospinal fluid (CSF) flow components modeled,
    motile cilia and cardiac pulsations, and their contributions to the flow patterns.
    \textbf{b.} Particle tracking data visualizing flow directionality in the middle ventricle.
    \textbf{c.} Flow velocity magnitudes in the middle ventricle from particle tracking data.
    \textbf{d.} Computational mesh with marked facets indicating regions of motile cilia (magenta) and
    anterior/posterior facets (blue).
    \textbf{e.} Streamlines of the (steady-state) CSF flow field simulated with the
    cilia-only/no-cardiac flow model.
    \textbf{f.} Streamlines of the CSF flow field simulated with the cardiac-only/no-cilia flow
    model at a time 25\% into the cardiac cycle. %t=0.112613 s, period=0.45045 s
    \textbf{g.} Streamlines simulated with the baseline flow model at
    a time 50\% into the cardiac cycle. %t=0.225225 s
    \textbf{h.} Streamlines simulated with the baseline flow model at
    a time 70\% into the cardiac cycle. %t=0.315315 s
    }
    \label{fig:fig1}
\end{figure}

@@

\subsection*{Computational CSF dynamics}
The beating motion of the cilia generates CSF flow with steady rotational
structures~\cite{Olstad2019CiliaryDevelopment}
within the brain ventricles (\Cref{fig:fig1}a--c).
In addition, cardiac pulsations induces pulsatile CSF flow.
We model this flow of CSF in the brain ventricles by the time-dependent,
incompressible Stokes equations which read as follows.
Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$
and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,\\
      \nabla \cdot \uu &= 0
      \quad  \text{in } \Omega,
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$
and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. 
Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions.
Furthermore, $\rho=1000 \ \mathrm{kg/m^3}$ is the CSF density
and $\mu=6.97\times 10^{-4} \ \mathrm{Pa\cdot s}$ is the dynamic viscosity
of the CSF~\cite{Bloomfield1998EffectsFluid}.
We introduce the tangential traction $\bsigpar$ as the
tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$,
and $P_{\nn}(\rr)$ is the tangential projection of a vector $\rr$ onto $\Gamma$.
We consider three types of boundary conditions:
(i) \emph{tangential traction} (no flow normal to the boundary
$\uu \cdot \nn = 0$ with a force $\btau$ applied tangentially on the boundary $\bsigpar = \btau$),
(ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in
the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$),
and (iii) \emph{free slip} (no flow normal to the boundary
$\uu \cdot \nn = 0$ and no tangential forces $\bsigpar = {\bf 0}$).
Unless noted otherwise, the free-slip condition is applied on the
boundary. We assume that the system starts at rest: $\uu(\xx, t=0) = \mathbf{0}$.

To study the impact of the motile cilia and the cardiac pulsations on the CSF flow,
we consider the following computational flow scenarios,
each predicting a CSF flow velocity $\uu$ and CSF pressure $p$.
\begin{itemize}
    \item A \emph{baseline} flow model representing a wild-type zebrafish including 
    contributions both from the cilia forces and the cardiac pulsations.
    In this scenario, we apply the tangential traction $\btau$ on the cilia boundary
    $\Gc$ and prescribe the normal pressure $\Tilde{p}$ on the anterior/posterior boundary $\Gp$. % Model 0/Model C

    \item A \emph{cilia-only/no-cardiac} flow model including contributions from the
    cilia forces but no cardiac pulsations.
    Here, we apply the tangential traction $\btau$ on the cilia boundary $\Gc$.
    In the absence of cardiac flow, the anterior/posterior boundary $\Gp=\emptyset$. % Model I/Model A

    \item A \emph{cardiac-only/no-cilia} flow model including contributions from the
    cardiac pulsations but no cilia forces.
    Here, we set the tangential traction to zero on the cilia boundary
    $\Gc$, while prescribing the normal pressure $\Tilde{p}$ on the
    anterior/posterior boundary $\Gp$.  % Model II/Model B
\end{itemize}

\subsection*{Cilia-driven flow}
On the ciliated regions of the ventricle walls $\Gc$ (\Cref{fig:fig1}d, magenta markers),
we impose a constant-in-time tangential traction $\bsigpar=\btau$ to represent the net forces
of the cilia acting on the CSF.
We set $\btau(\xx, \rr) = \tau \lambda(\xx) P_{\nn}(\rr)$, where the function $\lambda(\xx)$
and the sign of the vector $\rr=\pm(1, 0, 1)$ varies depending on the
cilia region (see~\Cref{subsec:appendixA1}).
Both $\rr$ and the value $\tau=6.5\times 10^{-4}$ Pa were chosen by 
model calibration using experimental data~\cite{Olstad2019CiliaryDevelopment}.
The data indicate speeds proximate to the ventricular walls of $27.4 \pm 5.4$ \textmu m/s and
$2.6 \pm 0.6$ \textmu m/s in the dorsal and ventral regions of the middle ventricle, respectively.
In the anterior ventricle, the speeds were estimated at $4.7 \pm 1.3$ \textmu m/s.
We calibrated $\tau$ heuristically by comparing these experimental velocity values
with the simulated, maximum velocity magnitude.

\subsection*{Flow induced by cardiac pulsations}\label{subsec:pressure_bcs}
To drive a cardiac-induced pulsatile flow inside the brain ventricles, we set $\Tilde{p}(t)=0$ and
$\Tilde{p}(t)=-A\sin{\omega t}$ on anterior and posterior parts, respectively, of the
boundary $\Gp$~(\Cref{fig:fig1}d, blue markers).
Based on a measured cardiac frequency of $f=2.22$ Hz~\cite{Olstad2019CiliaryDevelopment},
the angular cardiac frequency $\omega=2\pi f=6.97$ rad/s.
The amplitude $A=1.5\times 10^{-3}$ Pa determines the magnitude of the pulsatile flow,
and was calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}.
Similar to $\tau$, the amplitude $A$ was chosen by comparing experimental and simulated velocities.
Particle tracking data of pulsatile motion was used to determine the contributions of the
cardiac pulsations to the velocity magnitude,
as compared to the effect of cilia motion on the flow (see~\Cref{sec:appC_particle_transport}).

@@

\subsection*{Generation of \emph{Tg(5xuas:Signal-Dendra2)\textsuperscript{nw21Tg}} line}
\fixme{@Nathalie: This needs a more understandable title for us non-experts, could we change to
e.g. "Generation of transgenic line"?.}
The open reading frame (ORF) of the neuropeptide y (npy) signal peptide
(\emph{npy}: ENSDARG00000036222, Q1LW93) was fused with the N-terminal
of zebrafish codon optimized Dendra2 DNA sequence.
The DNA sequence was synthesized with EcoR1 enzyme sites at the 5’ and
3’ ends (GenScript Biotech), and inserted into the
pT2MUASMCS vector~\cite{Asakawa2008GeneticZebrafish}
through restriction enzyme cloning (GenScript Biotech).

@@

\subsection*{Intraventricular injection of microbeads and particle tracking}
To recover the flow fields shown in~\Cref{fig:fig1}b, c, we carried out
particle tracking in a larval zebrafish injected with fluorescent beads.
As described in Olstad~\emph{et al.}~\cite{Olstad2019CiliaryDevelopment},
anesthetized 2 dpf zebrafish were injected with 1 nl of a mixture containing 0.1\% w/v fluorescent beads
(SPHERO Fluorescent Yellow Particles 1\% w/v, F = 0.16 mm) diluted
in 7.5 mg/ml 70 kDa rhodamine B isothiocyanate-dextran (RITCdextran; Sigma-Aldrich, R9379)
dissolved in artificial CSF. Artificial CSF composition was as follows:
124 mM NaCl, 22 mM D-(+)-Glucose, 2.0 mM KCl, 1.6 mM MgSO\textsubscript{4} $\cdot$ 7 H\textsubscript{2}O,
1.3 mM KH\textsubscript{2}PO\textsubscript{4}, 24 mM NaHCO\textsubscript{3},
2.0 mM CaCl\textsubscript{2} $\cdot$ 2 H\textsubscript{2}O.
The needles used for the injections were pulled with a
Sutter Instrument Co. Model P-2000 from thin-walled glass
capillaries (1.00 mm; VWR) and cut open with a forceps.
A volume of 1 nl of the solution was injected with a pressure injector
(Eppendorf Femtojet 4i) in the rostral rhombencephalic ventricle.
The pressure and time were calibrated for each needle using a 0.01 mm calibration slide.

Following injection, the zebrafish was directly placed under the confocal
microscope and 1200 images were acquired at a frequency of 13.1 Hz.
A single optical section was obtained. Particle tracking was done using
TrackMate~\cite{Tinevez2017TrackMate:Tracking} in Fiji/ImageJ~\cite{Schindelin2012Fiji:Analysis}
and plotted using MATLAB, as described in Jeong~\emph{et al.}~\cite{Jeong2022MeasurementTelencephalon}.
The parameters for TrackMate were as follows.
The DoG detector was used with the settings: threshold: 2.0, with
median filtering; radius: 2.0, with subpixel localization.
Next, the Simple LAP tracker was used with settings: max frame gap: 2;
alternative linking cost factor: 1.05; linking max distance: 5.0; gap closing max distance: 2.0;
splitting max distance: 15.0; allow gap closing: true; allow track splitting:
false; allow track merging: false; merging max distance: 15.0; cutoff percentile: 0.9.
Only the particle tracks with at least 30 data points, while simultaneously
covering a distance of at least 8.8 $\mu$m, were plotted in~\Cref{fig:fig1}b, c.

@@

We seek to model the compartimentalized and pulsatile CSF flow observed \emph{in vivo}~(\Cref{fig:fig1}b--c).
To characterize CSF flow patterns predicted by
computational fluid dynamics within the brain ventricles, we begin by
considering the simulated CSF velocity and pressure when only considering
cilia motion, disregarding cardiac pressure pulsations. The
bidirectional traction induced by the motile cilia
inherently leads to a partial compartmentalization of the ventricular
system, with large-scale vortex structures in the anterior and middle
ventricles~(\Cref{fig:fig1}e). The highest velocities ($26.5$ \textmu m/s)
occur in the vicinity of the dorsal cilia in the middle ventricle.

To examine the effect of cardiac pressure pulsations,
we next consider a no-cilia model variation. In this scenario, the simulated CSF
flow is pulsatile as a result of the sinusoidal pressure gradient (\Cref{fig:fig1}g),
with a peak velocity magnitude of $4.8$ \textmu m/s. At each time instant, the flow
is unidirectional, with CSF initially flowing rostrocaudally (positive $x-$axis) before changing
direction to caudorostral in the middle of the cardiac cycle.

We now consider the baseline model, simulating flow driven by both cilia
and cardiac pulsations. We observe persistent flow features,
with vortex-structures in the anterior and middle ventricles
and directional flow in the posterior ventricle (\Cref{fig:fig1}e,
\fixme{Supplementary video 1}). These characteristic patterns suggest
that the flow in the anterior and middle ventricles are dominated by
the cilia contribution, while the cardiac pressure pulsations dominate
within the posterior ventricle. Over one cardiac cycle, the CSF flow
speed reaches up to 27.7 \textmu m/s, dorsally in the posterior part
of the middle ventricle. This peak is attained at the time instant
coinciding with the time of peak caudal cardiac pulsatile flow, as a
result of both the traction induced by the cilia and the pulsatile
pressure-driven flow being aligned with the rostrocaudal 
axis. Using the observed maximum velocity magnitude and choosing the height of
the middle ventricle ($110$ \textmu m) as a characteristic length
scale, yields a Reynolds number $\mathrm{Re} \approx 0.004$,
justifying the assumption of Stokes flow. In summary, the computational
CSF flow model reproduces the flow features observed \emph{in vivo}.

@@

To quantify the transport of solute, we plot the profile of the mean
concentrations $\cbar_i(t)$ as functions of time
(\Cref{fig:fig2}\fixme{b--d}) and
the times $\hat{t}$ when the mean concentrations $\cbar_i$ first exceed
a threshold value $\hat{c}$ (\Cref{fig:fig2}\fixme{f})
in 6 regions of interest (ROIs) (\Cref{fig:fig2}\fixme{e}).
The dynamics at the site of photoconversion (ROI 1)
is logarithmic and follows from imposing \eqref{eq:photoconversion_curve}.
For ROIs 2--6, we observe faster kinetics with higher diffusion coefficients
(\Cref{fig:fig2}\fixme{b--d}),
which results in smaller time to threshold
(\Cref{fig:fig2}\fixme{f}).
For all three diffusion coefficients, the dynamics of ROI 5
is significantly delayed compared to ROI 2--4,
and ROI 6 is the slowest. Notably, for $D_1$ the mean concentration
in ROI 6 is close to zero at the end of simulation.
To quantify the balance of advection versus diffusion as transport mehcanisms,
we calculate the Péclet numbers $\mathrm{Pe}_1=519$,
$\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for $D_1$, $D_2$ and $D_3$, respectively.
Since $\mathrm{Pe}>1$, this indicates that advection is the
dominant transport mechanism on a \fixme{global} scale.

@@

\subsection*{Computational model assumptions and boundary conditions}
To model CSF, we employed slip boundary conditions at the ventricular walls,
contrasting the classical choice of a noslip boundary condition.
The use of slip boundary conditions is motivated by experimental observations of the
flow field in zebrafish embryo brain ventricles~\cite{Olstad2019CiliaryDevelopment}.
It has been observed, both previously and in this work, that the fluid velocity increases towards the wall~(\Cref{fig:fig1}c).
Using slip boundary conditions allows for this behavior.
A couple of previous continuum model simulation studies of cilia-mediated flow~\cite{Siyahhan2014FlowVentricles, Thouvenin2020OriginCanal}
have modeled cilia as body forces. We chose to model the cilia as a force exerted tangentially on the boundary.
First, we deem this a good representation of the real nature of cilia,
since they are in reality microscale structures confined to the ventricular walls that
generate flow through collective motion. Second, it avoids the need of excessive mesh
refinement near the ciliated boundaries in the case that a
body force were to be applied in a $\sim$ 5 $\mu$m~\cite{Salman2022ComputationalEmbryo} deep cilia layer.

@@

@@

\textbf{Only-cilia/no-cardiac model.} This model version has cilia forces
as the only mechanism driving flow. We consider the
Stokes equations~\eqref{eq:appendix_stokes_eqs} with the slip boundary conditions
\begin{subequations}
    \begin{alignat}{2}
        \uu\cdot\nn &= 0 \quad &&\mathrm{on} \ \Gamma=\Gc\cup\Gs, \label{eq:modelA_slip_BC_impermeability} \\
        \bsigpar &= \btau \quad &&\mathrm{on} \ \Gc,
        \label{eq:modelA_slip_BC_shear} \\
        \bsigpar &= \mathbf{0} \quad &&\mathrm{on} \ \Gs,\label{eq:modelA_full_slip_BC}
    \end{alignat}\label{eq:modelA_slip_BC}%
\end{subequations}%
on the boundary $\Gamma=\Gamma_c\cup\Gamma_s$. The boundary $\Gamma_s$ is a free-slip boundary,
where we impose the free-slip condition~\eqref{eq:modelA_full_slip_BC}.
The first condition~\eqref{eq:modelA_slip_BC_impermeability} is the
impermeability condition, enforcing no flow normal to the boundary,
whereas~\eqref{eq:modelA_slip_BC_shear} imposes a stress tangential to
the boundary. These tangential stresses are applied to represent
the forces that the collective cilia motion exerts on the cerebrospinal fluid.

@@

in the weak form~\eqref{eq:modelA_weak_form1} of the Stokes equations.
We decompose the traction vector $\hat{\bsig}$ into a normal and tangential component:

@@

Inserting the above into~\eqref{eq:boundary_integral_BC1},
combined with application of the tangential traction
condition~\eqref{eq:modelA_slip_BC_shear} and the
free-slip condition~\eqref{eq:modelA_full_slip_BC}, yields
